Spoken Corpus of Karel Mako\v{n}

### 2018-04-12 2018-clarin-pisa/makon-data.tex
some blurb and data for clarin 2018

@@

\begin{itemize}
\item{interpretation of some passages in the New Testament, notably the parable
of the prodigal son, the parable of the talents, and the Our Father prayer,}
\item{milestone personal experiences, like the one in the concentration camp,}
\item{explaining symbolism, the apostles as symbols for human abilities,}
\item{references to specific people, like St. Teresa of \'{A}vila or Padre Pio.}
\end{itemize}

@@

We can take the current transcription %and the few notes some listeners made
%about topics and their alignment with the audio
as a source for a baseline.
Topic is inherently a vague concept, which I shall address when I focus on it
but for the baseline evaluation, we can take a look at some easily defined cases
like named entities. The ease about them is that we can look for their specific
word form and if it is not found, then it means the transcription is not talking
about it.

Of course, there are exceptions, like saying ``the capital of Japan'' instead of
``Tokyo'' but firstly, I chose terms where this is not likely to happen and
secondly, this can only lower the recall of our baseline score, not the
precision, and we have no way of telling the precision given no annotated data
for this task.\footnote{The 50-minute test set is way too small for this kind of
task.}

I have chosen the following terms:

@@

@@

Sometimes the quality changes abruptly inside a file, when recording was stopped
and resumed with different conditions.

Expectably, the age of a recording correlates with its quality but other factors
play a significant role also. Apart of obvious ones like the relative position
of the speaker to the microphone, the acoustics of the room and quality of the
recording device, also the tape brand seems to be quite important.

>>>>

Most of the previous work was focused on acquiring the automatic transcription
and development of a web interface to allow users to access the talks and
provide corrections to the existing transcription, in order to have high-quality
transcription as well as to gather more training data.

The actual content of the corpus with respect to topics, references and
statements, is on one hand quite well known firstly because there are many books
by the author with descriptive indexes like any other book, which largely
correspond to what is said in the talks, secondly because the the domain stays
more or less consistent across the whole set. On the other hand, it is only
known vaguely and a systematic effort to analyze it is yet to be carried out.

@@

\begin{itemize}
\item{interpretation of some passages in the New Testament, notably the parable
of the prodigal son, the parable of the talents, and the Our Father prayer,}
\item{milestone personal experiences, like the one in the concentration camp,}
\item{explaining symbolism, like the apostles as symbols for human abilities,}
\item{references to specific people, like St. Teresa of \'{A}vila or Padre Pio.}
\end{itemize}

@@

We can take the current transcription %and the few notes some listeners made
%about topics and their alignment with the audio
as a source for a baseline.
Topic is inherently a vague concept %, which I shall address when I focus on it
but for the baseline evaluation, we can take a look at some easily defined cases
like named entities. The ease about them is that we can look for their specific
word form and if it is not found, then it means the transcription is not talking
about it.

Of course, there are exceptions, like saying ``the capital of Japan'' instead of
``Tokyo'' but firstly, I chose terms where this is not likely to happen and
secondly, this can only lower the recall of our baseline score, not the
precision, and we have no way of telling the recall given no annotated data
for this task.\footnote{The 50-minute test set is way too small for this kind of
task. Another source could be manual notes available to some of the recordings
but preparing them for this purpose requires some more effort.}

I have chosen the following example topics:

@@

The precision of over 80\% seems quite promising and hints that the current
quality of transcription allows for reasonable searching through the corpus.
However, the hits were all clearly audible. I met none in an acoustically defect
(see the next subsection) file, which leads to the hypothesis that those are not
yet searchable to a distantly satisfying level.

\subsubsection{Correspondence between Talks and Books}

The proclamation at the end of the introduction~\ref{intro}, that the
topics of the talks correspond largely to that of the concurrently written
books, deserves a kind of support. Mako\v{n} translated the book \emph{El
Castillo Interior}, English \emph{The Interior Castle} by St. Teresa of Ávila in
1977. For most recordings, the year of their acquisition is known.
Figure~\ref{fig:teresa-year} shows how many hits for the name of the Saint are
present in the transcription by recording year.

@@

Sometimes the quality changes abruptly inside a file, when recording was stopped
and resumed with different conditions. This prevents us from classifying
individual recordings by their acoustic quality.

Expectably, the age of a recording correlates with its quality but other factors
play a significant role also. Most are obvious like the relative position
of the speaker to the microphone, the acoustics of the room, quality of the
recording device and also the tape brand. Unfortunately, none of these can be
said with certainty.
